---
title: "Data Visualization"
author: "Vinh"
output:
  html_document:
    df_print: paged
    code_folding: hide
    toc: true
    toc_float: 
      collapsed: false
      smooth_scroll: false
    toc_depth: '2'
    css: styles.css
---
# Visualize amount
## Dot plot/lollipop chart

```{r warning = FALSE, message = FALSE, results = FALSE}
library(tidyverse)
library(ggforce)
library(gapminder)
library(ggridges)
library(PASWR)
library(tinter)
library(fmsb)
library(ggalt)
library(cowplot)
library(PASWR)
library(RColorBrewer)
library(Stat2Data)
library(colorspace)
library(reshape2)
library(devtools)
library(waffle)
library(treemapify)
library(zoo)
library(wordcloud2)
library(tm) #text mining

text_color = "#353d42"
caption_color = "#666666"
font = "Avenir Next"
```

### One group
```{r warning = FALSE, message = FALSE, out.width='100%', fig.asp= 0.8, fig.align='center'}
gapminder_dot <- gapminder %>% 
  filter(year == 2007, continent == "Americas") %>%
  dplyr::select(country, lifeExp) %>% 
  mutate(country = fct_reorder(country, lifeExp),
         mean_life = mean(lifeExp),
         flag  = ifelse(lifeExp - mean_life > 0, TRUE, FALSE))

ggplot(gapminder_dot, aes(x = lifeExp, y = country)) +
  geom_segment(
    aes(x = 60, xend = lifeExp, y = country, yend = country),
    color = "#0072B2",
    size = 0.5,
    alpha = 0.3
  ) +
  geom_point(
    color = "#0072B2", 
    size = 3) +
  scale_x_continuous(
    name = "life expectancy (years)",
    expand = c(0, 0), 
    limits = c(59.7, 81.5)) +
  scale_y_discrete(
    name = NULL, 
    expand = c(0, 0.5)) +
  labs(
    caption = "Source: gapminder",
    title = "Life expectancy of American countries in 2007") +
  theme(
    axis.ticks = element_blank(),
    axis.text = element_text(family = "Avenir Next", size = 11, color = "#353d42"),
    panel.background = element_blank(),
    panel.grid.major = element_line(color = "#cbcbcb", size = 0.1),
    plot.margin = margin(18, 6, 3, 1.5),
    axis.title.x = element_text(family = "Avenir Next", size = 11, color = "#353d42", margin = margin (t = 10)),
    plot.title = element_text(family = "Avenir Next", color = "#353d42", size = 15, face = "bold", margin = margin(b = 20)),
    plot.caption = element_text(family = "Avenir Next", color = "#666666", size = 9, hjust = 0, vjust = 2))

```

### Change baseline
```{r warning = FALSE, message = FALSE, out.width='100%', fig.asp= 0.8, fig.align='center'}

ggplot(gapminder_dot, aes(x = lifeExp, y = country, color = flag)) +
  geom_segment(
    aes(x = mean_life , xend = lifeExp, y = country, yend = country),
    size = 0.5,
    alpha = 0.3
  ) +
  geom_point(
    size = 3) +
  annotate("text", x = 81, y = 14.5, 
           label = "Above average", 
           family = "Avenir Next",
           color = "#353d42", 
           size = 3) +
  annotate("text", x = 69, y = 10, 
           label = "Below average",
           family = "Avenir Next",
           color = "#353d42", 
           size = 3) +
  geom_curve(aes(x = 69, xend = 71, y = 9, yend = 7),
             color = "#353d42",
             size = 0.1,
             arrow = arrow(length = unit(0.01, "npc"))) +
  geom_curve(aes(x = 81, xend = 79, y = 15, yend = 17),
             color = "#353d42",
             size = 0.1,
             arrow = arrow(length = unit(0.01, "npc"))) +
  scale_x_continuous(
    name = "life expectancy (years)",
    expand = c(0, 0), 
    limits = c(59.7, 85)) +
  scale_y_discrete(
    name = NULL, 
    expand = c(0, 0.5)) +
   scale_color_manual(
    values = c("#D55E00", "#0072B2")
  ) +
  labs(
    caption = "Source: gapminder",
    title = "Life expectancy of American countries in 2007") +
  theme(
    axis.ticks = element_blank(),
    axis.text = element_text(family = "Avenir Next", size = 11, color = "#353d42"),
    panel.background = element_blank(),
    panel.grid.major = element_line(color = "#cbcbcb", size = 0.1),
    plot.margin = margin(18, 6, 3, 1.5),
    axis.title.x = element_text(family = "Avenir Next", size = 11, color = "#353d42", margin = margin (t = 10)),
    legend.position = "none",
    plot.title = element_text(family = "Avenir Next", color = "#353d42", size = 15, face = "bold", margin = margin(b = 20)),
    plot.caption = element_text(family = "Avenir Next", color = "#666666", size = 9, hjust = 0, vjust = 2))

```

## Dumbbell chart 
```{r warning = FALSE, message = FALSE, results = FALSE}
gapminder_dot_1 <- gapminder %>% 
  filter(year %in% c(1987, 2007), continent == "Americas") %>%
  dplyr::select(year, country, lifeExp) %>% 
  mutate(country = fct_reorder(country, lifeExp)) %>% 
  spread(year, lifeExp)

ggplot(gapminder_dot_1) +
  geom_segment(
    aes(x = `1987`, xend = `2007`, y = country, yend = country),
    color = "black",
    alpha = 0.3
  ) +
  geom_point(
    aes(x = `1987`, y = country),
    color = "#D55E00", 
    size = 2) +
  geom_point(
    aes(x = `2007`, y = country),
    color = "#0072B2", 
    size = 2) +
  scale_x_continuous(
    name = "life expectancy (years)",
    expand = c(0, 0), 
    limits = c(52, 81.5)) +
  scale_y_discrete(
    name = NULL, 
    expand = c(0, 0.5)) +
  labs(
    caption = "Source: gapminder",
    title = "Life expectancy of American countries 1987 - 2007") +
  theme(
    axis.ticks = element_blank(),
    axis.text = element_text(family = "Avenir Next", size = 11, color = "#353d42"),
    panel.background = element_blank(),
    panel.grid.major = element_line(color = "#cbcbcb", size = 0.1),
    plot.margin = margin(18, 6, 3, 1.5),
    axis.title.x = element_text(family = "Avenir Next", size = 11, color = "#353d42", margin = margin (t = 10)),
    plot.title = element_text(family = "Avenir Next", color = "#353d42", size = 15, face = "bold", margin = margin(b = 20)),
    plot.caption = element_text(family = "Avenir Next", color = "#666666", size = 9, hjust = 0, vjust = 2))
```

```{r out.width='100%', fig.asp= 0.8, fig.align='center', val = TRUE}
# Another method use geom_dumbbell from library ggalt

dumbbell_df <- gapminder %>%
  filter(year == 1967 | year == 2007) %>%
  dplyr::select(country, year, lifeExp) %>%
  spread(year, lifeExp) %>%
  mutate(gap = `2007` - `1967`) %>%
  arrange(desc(gap)) %>%
  head(10)

#Make plot
ggplot(dumbbell_df, aes(x = `1967`, xend = `2007`, y = reorder(country, gap), group = country)) + 
  geom_dumbbell(colour = "#dddddd",
                size = 3,
                colour_x = "#FAAB18",
                colour_xend = "#1380A1") +
  labs(title="Life expectancy rise in 10 countries, 1967-2007 ",
       subtitle="We are living longer",
       caption = "Source: gapminder") +
  theme(
    panel.background = element_blank(),
    panel.grid.major.y = element_line(color = "#d9dbda", size = 0.4),
    axis.ticks = element_blank(),
    axis.text = element_text(family = "Avenir Next", size = 11, color = "#353d42"),
    axis.title = element_blank(),
    plot.title = element_text(family = "Avenir Next", size = 15, color = "#353d42", face = "bold"),
    plot.subtitle = element_text(family = "Avenir Next", size = 11, color = "#353d42"),
    plot.caption = element_text(family = "Avenir Next", size = 10, color = "#666666", hjust = 0, margin = margin(t = 10))
  )
```

## Radar chart

```{r warning = FALSE, message = FALSE, results = FALSE}

exam_scores <- data.frame(
    row.names = c("Student.1", "Student.2", "Student.3"),
      Biology = c(7.9, 3.9, 9.4),
      Physics = c(10, 20, 0),
        Maths = c(3.7, 11.5, 2.5),
        Sport = c(8.7, 20, 4),
      English = c(7.9, 7.2, 12.4),
    Geography = c(6.4, 10.5, 6.5),
          Art = c(2.4, 0.2, 9.8),
  Programming = c(0, 0, 20),
        Music = c(20, 20, 20)
)
# To use the fmsb package, the data should be organized as follow:
# - row 1 contains maximum values for each variable
# - row 2 contains minimum values for each variable
# - data for cases or individuals start from row 3
# - number of columns must be more than 2

max_min <- data.frame(
  Biology = c(20, 0), Physics = c(20, 0), Maths = c(20, 0),
  Sport = c(20, 0), English = c(20, 0), Geography = c(20, 0),
  Art = c(20, 0), Programming = c(20, 0), Music = c(20, 0)
)
rownames(max_min) <- c("Max", "Min")

# Bind the variable ranges to the data
(df <- rbind(max_min, exam_scores))

```

### One observation
```{r out.width='100%', fig.asp= 1, fig.align='center'}
# rada chart of student 1
student1_data <- df[c("Max", "Min", "Student.1"), ]

radarchart(student1_data,
           axistype = 1,
           # Customize the polygon
           pcol = "#1b9e77", #color of the line
           pfcol = scales::alpha("#1b9e77", 0.5), # filling color
           plwd = 1.5, # line width
           # Customize the grid
           cglcol = "#cbcbcb", #color
           cglwd = 0.9, #line width
           cglty = 3,
           axislabcol = "#cbcbcb",
           title = "Student A score report",
           vlabels = colnames(student1_data),
           vlcex = 0.7,
           caxislabels = c(0, 5, 10, 15, 20)
           )
```

### Multiple observations 1
```{r out.width='100%', fig.asp= 1, fig.align='center'}
# rada chart of student 1
radarchart(df,
           axistype = 1,
           # Customize the polygon
           pcol = c("#e41a1c", "#377eb8", "#4daf4a"), #color of the line
           pfcol = scales::alpha(c("#e41a1c", "#377eb8", "#4daf4a"), 0.6), # filling color
           plwd = c(1.5, 1.5, 1.5), # line width
           plty = c(1, 1, 1),
           # Customize the grid
           cglcol = "#cbcbcb", #color
           cglwd = 0.9, #line width
           cglty = 3,
           axislabcol = "#cbcbcb",
           title = "Score report of three students ",
           vlabels = colnames(student1_data),
           vlcex = 0.7,
           caxislabels = c(0, 5, 10, 15, 20)
           )
```

### Multiple observations 2
```{r out.width='100%', fig.asp= 0.5, fig.align='center'}

# Define colors and titles
colors <- c("#00AFBB", "#E7B800", "#FC4E07")
titles <- c("Student A", "Student B", "Student C")

# Reduce plot margin using par()
# Split the screen in 3 parts
par(mfrow = c(1,3))

# Create the radar chart
for(i in 1:3){
  radarchart(df[c(1, 2, i+2), ],
           axistype = 1,
           # Customize the polygon
           pcol = colors[i], #color of the line
           pfcol = scales::alpha(colors[i], 0.5), # filling color
           plwd = 1.5, # line width
           # Customize the grid
           cglcol = "#cbcbcb", #color
           cglwd = 0.9, #line width
           cglty = 3,
           axislabcol = "#cbcbcb",
           title = titles[i],
           vlabels = colnames(df[c(1, 2, i+2), ]),
           vlcex = 0.7,
           caxislabels = c(0, 5, 10, 15, 20)
           )
}
```

### Compared with the average
```{r out.width='100%', fig.asp= 0.4, fig.align='center', message = FALSE, warning = FALSE, results = FALSE}

set.seed(123)
df <- as.data.frame(
  matrix(sample(2:20 , 90 , replace = TRUE),
         ncol=9, byrow = TRUE)
  )
colnames(df) <- c(
  "Biology", "Physics", "Maths", "Sport", "English", 
  "Geography", "Art", "Programming", "Music"
  )
rownames(df) <- paste0("Student.", 1:nrow(df))
head(df)

# Rescale each variable to range between 0 and 1
library(scales)
df_scaled <- round(apply(df, 2, scales::rescale), 2)
df_scaled <- as.data.frame(df_scaled)
head(df_scaled)

#Prepare data
# Variables summary
# Get the minimum and the max of every column  
col_max <- apply(df_scaled, 2, max)
col_min <- apply(df_scaled, 2, min)
# Calculate the average profile 
col_mean <- apply(df_scaled, 2, mean)
# Put together the summary of columns
col_summary <- t(data.frame(Max = col_max, Min = col_min, Average = col_mean))


# Bind variables summary to the data
df_scaled2 <- as.data.frame(rbind(col_summary, df_scaled))
head(df_scaled2)
```

```{r out.width='100%', fig.asp= 0.8, fig.align='center', message = FALSE, warning = FALSE}
opar <- par() 
# Define settings for plotting in a 3x4 grid, with appropriate margins:
par(mar = rep(0.8,4))
par(mfrow = c(3,4))
# Produce a radar-chart for each student
for (i in 4:nrow(df_scaled2)) {
  radarchart(
    df_scaled2[c(1:3, i), ],
    pfcol = c("#99999980",NA),
    pcol= c(NA,2), plty = 1, plwd = 2,
    title = row.names(df_scaled2)[i]
  )
}
# Restore the standard par() settings
par <- par(opar) 
```

## Bar plot

### Vertical bars

```{r message = FALSE, out.width='100%', fig.asp=.75, fig.align='center'}
boxoffice <- data.frame(rank = 1:5,
                title = c("Star Wars: The Last Jedi", "Jumanji: Welcome to the Jungle", "Pitch Perfect 3", "The Greatest Showman", "Ferdinand"),
                title_short = c("Star Wars", "Jumanji", "Pitch Perfect 3", "Greatest Showman", "Ferdinand"),
                amount = c(71565498, 36169328, 19928525, 8805843, 7316746),
                amount_text = c("$71,565,498", "$36,169,328", "$19,928,525", "$8,805,843", "$7,316,746"))

ggplot(boxoffice, aes(x = fct_reorder(title_short, rank), y = amount)) +
    geom_col(
      fill = "#56B4E9", 
      width = 0.6, 
      alpha = 0.9)+
    scale_y_continuous(
      expand = c(0,0),
      breaks = c(0, 2e7, 4e7, 6e7),
      labels = c("0", "20", "40", "60"),
      name = "Weekend gross (million USD)") +
    scale_x_discrete(
      name = NULL,
      expand = c(0, 0.4)) +
    coord_cartesian(clip = "off") +
    labs(
      caption = "Source: Box Office Mojo",
      title = "Highest grossing movies for 22-24.12.2017") +
    theme(
      axis.ticks = element_blank(),
      axis.text = element_text(family = font, color = text_color, size = 11),
      axis.title.y = element_text(family = font, color = text_color, size = 11, margin = margin(r = 20)),
      panel.background = element_blank(),
      panel.grid.major.y = element_line(color = "#cbcbcb", size = 0.5),
      plot.title = element_text(family = font, size = 15, color = text_color, face = "bold", margin = margin(b = 10), hjust = 0),
      plot.caption = element_text(family = font, size = 10, color = "#666666", margin = margin(t = 15), hjust = 0)
    )
```

### Horizontal bars

```{r message = FALSE, out.width='100%', fig.asp=.7, fig.align='center'}
ggplot(boxoffice, aes(x = fct_reorder(title_short, desc(rank)), y = amount)) +
    geom_col(
      fill = "#56B4E9", 
      alpha = 0.9)+
    scale_y_continuous(
      expand = c(0,0),
      breaks = c(0, 2e7, 4e7, 6e7),
      labels = c("0", "20", "40", "60"),
      name = "Weekend gross (million USD)") +
    scale_x_discrete(name = NULL) +
    coord_cartesian(clip = "off") +
    labs(
      title = "Highest grossing movies for 22-24.12.2017",
      caption = "Source: Box Office Mojo")  +
    coord_flip() +
    theme(
      axis.ticks = element_blank(),
      axis.text = element_text(family = font, color = "#353d42", size = 11),
      axis.text.x = element_text(vjust = 1, hjust = 1),
      axis.title = element_text(family = font, color = "#353d42", size = 11),
      panel.background = element_blank(),
      panel.grid.major.y = element_line(color = "#cbcbcb", size = 0.5),
      plot.title = element_text(family = font, size = 15, color = "#353d42", face = "bold", margin = margin(b = 10), vjust = -2),
      plot.caption = element_text(family = font, size = 10, color = "#666666", hjust = 0)    
      )
```

### Grouped bars

```{r message = FALSE, out.width='100%', fig.asp=.6, fig.align='center'}
diamonds_by_color_cut <- 
  diamonds %>% 
  group_by(color, cut) %>% 
  summarize(average_price = mean(price))

color_five = RColorBrewer::brewer.pal(n = 9, "BuGn")[4:8] #Get 5 color from the scale
#Focus on price differences between each color group
ggplot(diamonds_by_color_cut, aes(x = color, y = average_price, fill = cut)) +
  geom_col(position = "dodge", alpha = 0.9) +
  scale_y_continuous(
    expand = c(0, 0),
    name = "Average price (USD)",
    breaks = c(1000, 2000, 3000, 4000, 5000, 6000),
    labels = c("1,000", "2,000","3,000","4,000","5,000","6,000")) + 
  scale_fill_manual(values = color_five, name = NULL) + 
  coord_cartesian(clip = "off") +
  labs(
    title = "Prices of 50,000 diamonds versus cut quality and color",
    caption = "Source: diamonds - ggplot2") +
  xlab("Color") +
  theme(
    panel.background = element_blank(),
    panel.grid.major.y = element_line(color = "#cbcbcb", size = 0.5),
    axis.ticks = element_blank(),
    axis.text = element_text(family = font, color = text_color, size = 11),
    axis.title = element_text(family = font, color = text_color, size = 11),
    axis.title.x = element_text(margin = margin(t = 8)),
    axis.title.y = element_text(margin = margin(r = 8)),
    plot.caption = element_text(family = font, size = 10, color = text_color, hjust = 0),
    plot.title = element_text(family = font, size = 15, color = text_color, face = "bold"),
    legend.text = element_text(family = font, color = text_color, size = 8)
    )
```

### Stacked Bars

```{r message = FALSE, out.width='100%', fig.asp=.8, fig.align='center'}

titanic = read.csv("/Users/huvi/Downloads/Project_website/Dataset/titanic.csv")
titanic_stacked <- titanic %>% 
  group_by(Pclass, Sex) %>% 
  tally() %>% 
  arrange(Pclass, desc(Sex))%>% 
  mutate(Sex = factor(Sex, levels = c("female", "male"))) %>% 
  group_by(Pclass) %>% 
  mutate(nlabel = cumsum(n) - n/2, Pclass = paste(Pclass, "class"))

ggplot(titanic_stacked, aes(x = Pclass, y = n, fill = Sex)) +
  geom_col(position = "stack", color = "white", size = 1, width = 1) +
  geom_text(aes(y = nlabel, label = n), color = "white", size = 6, family = font) +
  labs( title = "Numbers of passengers on the Titanic based on gender and class",
        caption = "Source: Encyclopedia Titanica") +
  scale_x_discrete(expand = c(0,0), name = NULL) +
  scale_y_continuous(expand = c(0,0), name = NULL, breaks = NULL) +
  coord_cartesian(clip = "off") +
  scale_fill_manual(values = c("#D55E00", "#0072B2"),
                    breaks = c("female", "male"),
                    labels = c("male passengers", "female passengers"),
                    name = NULL) +
  theme(panel.background = element_blank(),
        axis.text = element_text(family = font, color = text_color, size = 11),
        axis.text.x = element_text(margin = margin(t = 15)),
        legend.text = element_text(family = font, color = text_color, size = 11),
        legend.position = "bottom",
        legend.justification = "center",
        legend.spacing.x = grid::unit(7, "pt"),
        legend.spacing.y = grid::unit(0, "cm"),
        plot.title = element_text(family = font, size = 15, color = text_color, face = "bold", margin = margin(b = 30)),
        plot.caption = element_text(family = font, size = 10, color = text_color, hjust = 0))
```

## Heatmap

```{r warning = FALSE, message = FALSE, out.width='100%', fig.asp= 0.8, fig.align='center'}

country_list = c("United States", "China", "India", "Japan", "Vietnam", "Brazil", "Germany", "France", "United Kingdom", "Italy", "New Zealand", "Canada", "Mexico", "Chile", "Argentina", "Norway", "South Africa", "Myanmar", "Israel", "Iceland")

internet <- read.csv("/Users/huvi/Downloads/Project_website/Dataset/Individuals.using.the.Internet.%.population.World.Bank.csv")

internet_short <- internet %>% 
  dplyr::select(Country = Country.Name, Time, Users = Individuals.using.the.Internet....of.population...IT.NET.USER.ZS.) %>% 
  filter(Country %in% country_list) %>% 
  mutate(Users = ifelse(is.na(Users), 0, Users))

internet_summary <- internet_short %>% 
  group_by(Country) %>% 
  summarize(Time1 = min(Time[Users > 0]),
            Last = Users[n()]) %>% 
  arrange(Last, desc(Time1))
         
internet_short <- internet_short %>% 
  mutate(Country = factor(Country, levels = internet_summary$Country),
         Users = as.numeric(Users))

ggplot(internet_short, aes(x = Time, y = Country, fill = Users)) +
  geom_tile(color = "white", size = 0.3) + #Color of border between tiles and size of the color
  scale_x_continuous(expand = c(0, 0), name = NULL) +
  scale_y_discrete(name = NULL, position = "right")+
  scale_fill_viridis_c(
    option = "A", begin = 0.05, end = 0.98,
    limits = c(0,100), #Set limit on the legend
    name = "internet users / 100 people",
    guide = guide_colorbar(
      direction = "horizontal",
      label.position = "bottom",
      title.position = "top",
      barwidth = grid::unit(2.5, "in"),
      barheight = grid::unit(0.2, "in"), 
      ticks = FALSE)) +
  labs(
    caption = "Source: World Bank",
    title = "Percentage of internet users in 20 countries between 1999 and 2019") +
  theme(
    panel.background = element_blank(),
    axis.line = element_blank(),
    axis.ticks.length = grid::unit(1, "pt"),
    axis.ticks = element_blank(),
    legend.position = "bottom",
    legend.justification = "right",
    legend.title = element_text(family = "Avenir Next", color = "#353D42", size = 11),
    legend.title.align = 0.5,
    legend.box.spacing = unit(0, "pt"),
    legend.text = element_text(family = "Avenir Next", color = "#353D42", size = 9),
    axis.text = element_text(family = "Avenir Next", color = "#353D42", size = 11),
    plot.title = element_text(family = "Avenir Next", color = "#353d42", size = 13, face = "bold", hjust = 0, margin = margin(b = 10)),
    plot.caption = element_text(family = "Avenir Next", color = "#666666", size = 9, hjust = 0))
```

## Word cloud

```{r warning = FALSE, message = FALSE, out.width='100%', fig.asp= 1.1, fig.align='center'}
# Read the text
HP <- readLines("/Users/huvi/Downloads/Project_website/Dataset/HP_Part1_Text.txt")
# Load as a corpus (a list of document, only 1 in this case)
docs <- Corpus(VectorSource(HP))

toSpace <- content_transformer(function (x , pattern ) gsub(pattern, " ", x))
#docs <- tm_map(docs, toSpace, "/")
#docs <- tm_map(docs, toSpace, "@")
#docs <- tm_map(docs, toSpace, "—")

# Convert the text to lower case
docs <- tm_map(docs, content_transformer(tolower))
# Remove numbers
docs <- tm_map(docs, removeNumbers)
# Remove english common stopwords
docs <- tm_map(docs, removeWords, stopwords("english"))
# Remove your own stop word
# specify your stopwords as a character vector
docs <- tm_map(docs, removeWords, c("blabla1", "blabla2")) 
# Remove punctuations
docs <- tm_map(docs, removePunctuation)
# Eliminate extra white spaces
docs <- tm_map(docs, stripWhitespace)
# Text stemming
# docs <- tm_map(docs, stemDocument)

dtm <- TermDocumentMatrix(docs)
m <- as.matrix(dtm)
v <- sort(rowSums(m),decreasing=TRUE)
d <- data.frame(word = names(v),freq=v)
d <- d[-c(2,4,11,15,28,34,42,43),]

wordcloud2(data = d, 
           size = 1.5, 
           color = "random-dark",
           fontFamily = "Avenir Next")
```

# Visualize distribution

```{r warning = FALSE, message = FALSE, results = FALSE}
titanic = read.csv("/Users/huvi/Downloads/Project_website/Dataset/titanic.csv")

age_counts <- hist(titanic$Age, breaks = (0:16) * 5 + .01, plot = FALSE)$counts
age_hist <- data.frame(
  "Age_range" = c(c("0--5", "6--10", "11--15", "16--20", "21--25", "26--30", "31--35", "36--40", "41--45", "46--50", "51--55", "56--60", "61--65", "66--70", "71--75"), "76--80"),
  Count = age_counts,
  check.names = FALSE
)
```

## Histogram
### Single histogram
```{r out.width='100%', fig.asp= 0.75, fig.align='center'}
age_hist = cbind(age_hist, age = (1:16) * 5 - 2.5)

ggplot(age_hist, aes(x = age, y = Count )) +
  geom_col(width = 4.7, fill = "#56B4E9") + 
  scale_y_continuous(expand = c(0,0), breaks = 25 * (0:6)) +
  scale_x_continuous(expand = c(0,0), name = "age(years)", limits = c(0,80), breaks = seq(0, 80, 10))+
  coord_cartesian(clip = "off") +
  labs(
    caption = "Source: Encyclopedia Titanica",
    title = "Number of passengers with known age on the Titanic") +
  theme(
    panel.background = element_blank(),
    panel.grid.major.y = element_line(color = "#cbcbcb", size = 0.5),
    axis.ticks = element_blank(),
    axis.text = element_text(family = "Avenir Next", color = "#353d42", size = 11),
    axis.title = element_text(family = "Avenir Next", color = "#353d42", size = 11),
    plot.title = element_text(family = "Avenir Next", size = 13, color = "#353d42", face = "bold", margin = margin(b = 10)),
    plot.caption = element_text(family = "Avenir Next", size = 10, color = "#666666", hjust = 0)
  )
```

```{r warning = FALSE, message = FALSE, results = FALSE}
data.frame(
  age = (1:16) * 5 - 2.5, 
  male = hist(filter(titanic, Sex == "male")$Age, breaks = (0:16) * 5 + .01, plot = FALSE)$counts,
  female = hist(filter(titanic, Sex == "female")$Age, breaks = (0:16) * 5 + .01, plot = FALSE)$counts) %>% 
  gather(gender, count, -age) -> gender_counts

gender_counts$gender <- factor(gender_counts$gender, levels = c("female", "male"))
```

### Two histograms
```{r out.width='100%', fig.asp= 0.75, fig.align='center',  eval = FALSE}
ggplot(gender_counts, aes(x = age, y = count, fill = gender)) +
  geom_col(position = "stack") + 
  scale_y_continuous(expand = c(0,0), breaks = 25 * (0:10)) +
  scale_x_continuous(expand = c(0,0), name = "age(years)", limits = c(0,80), breaks = seq(0, 80, 10))+
  coord_cartesian(clip = "off") +
  scale_fill_manual(values = c("#D55E00", "#0072B2")) +
  labs(
    caption = "Source: Encyclopedia Titanica",
    title = "Numbers of passenger with known age on the Titanic") +
  theme(
    panel.background = element_blank(),
    panel.grid.major.y = element_line(color = "#cbcbcb", size = 0.5),
    axis.ticks = element_blank(),
    axis.text = element_text(family = "Avenir Next", color = "#353d42", size = 1),
    axis.title = element_text(family = "Avenir Next", color = "#353d42", size = 11),
    legend.text = element_text(family = "Avenir Next", color = "#353d42", size = 11),
    legend.title = element_text(family = "Avenir Next", color = "#353d42", size = 11),
    legend.position = c(.9, .87),
    legend.justification = c("right", "top"),
    plot.title = element_text(family = "Avenir Next", size = 13, color = "#353d42", face = "bold", margin = margin(b = 10)),
    plot.caption = element_text(family = "Avenir Next", size = 10, color = "#666666", hjust = 0)
  )
```

```{r out.width='100%', fig.asp= 0.75, fig.align='center',  eval = FALSE}

ggplot(gender_counts, aes(x = age, y = count, fill = gender)) +
  geom_col(position = "identity", alpha = 0.7) + 
  scale_y_continuous(expand = c(0,0), breaks = 25 * (0:10)) +
  scale_x_continuous(expand = c(0,0), name = "age(years)", limits = c(0,80), breaks = seq(0, 80, 10))+
  coord_cartesian(clip = "off") +
  scale_fill_manual(
    values = c("#D55E00", "#0072B2"),
    guide = guide_legend(reverse = TRUE)) +
  labs(
    caption = "Source: Encyclopedia Titanica",
    title = "Numbers of passenger with known age on the Titanic") +
  theme(
    panel.background = element_blank(),
    panel.grid.major.y = element_line(color = "#cbcbcb", size = 0.5),
    axis.ticks = element_blank(),
    axis.text = element_text(family = "Avenir Next", color = "#353d42", size = 11),
    axis.title = element_text(family = "Avenir Next", color = "#353d42", size = 11),
    plot.title = element_text(family = "Avenir Next", size = 13, color = "#353d42", face = "bold", margin = margin(b = 10)),
    plot.caption = element_text(family = "Avenir Next", size = 10, color = "#666666", hjust = 0),
    legend.position = c(.9, .87),
    legend.justification = c("right", "top"),
    legend.text = element_text(family = "Avenir Next", color = "#353d42", size = 11),
    legend.title = element_text(family = "Avenir Next", color = "#353d42", size = 11)
  )
```

```{r out.width='100%', fig.asp= 0.75, fig.align='center', echo = FALSE}

ggplot(gender_counts, aes(x = age, y = ifelse(gender == "male", -1, 1)*count, fill = gender)) +
  geom_col() + 
  scale_y_continuous(
    expand = c(0,0), 
    breaks = 20 * (-5:3),
    name = "count",
    labels = c("100","80","60","40","20","0","20","40","60")) +
  scale_x_continuous(expand = c(0,0), name = "age(years)", limits = c(0,83), breaks = seq(0, 80, 20))+
  coord_flip() +
  scale_fill_manual(
    values = c("#D55E00", "#0072B2"),
    guide = "none") +
  labs(
    caption = "Source: Encyclopedia Titanica",
    title = "Numbers of passenger with known age on the Titanic") +
  draw_text(x = 70, y = -39, "male", hjust = 0, size = 11, family = "Avenir Next", color = "#353d42") +
  draw_text(x = 70, y = 21, "female", hjust = 0, size = 11, family = "Avenir Next", color = "#353d42") +
  theme(
    panel.background = element_blank(),
    panel.grid.major = element_line(color = "#cbcbcb", size = 0.2),
    axis.ticks = element_blank(),
    axis.text = element_text(family = "Avenir Next", color = "#353d42", size = 11),
    axis.title = element_text(family = "Avenir Next", color = "#353d42", size = 11),
    plot.title = element_text(family = "Avenir Next", size = 13, color = "#353d42", face = "bold", margin = margin(b = 10)),
    plot.caption = element_text(family = "Avenir Next", size = 10, color = "#666666", hjust = 0))
```

## Density plot
### Single density
```{r out.width='100%', fig.asp= 0.7, fig.align='center', echo = FALSE}
ggplot(titanic, aes(x = Age)) +
  geom_density_line(fill = "#56B4E9", color = "darkblue", alpha = 0.9, bw = 2, kernel = "gaussian") +
  scale_y_continuous(
    expand = c(0,0),
    limits = c(0, 0.045),
    name = "density") +
  scale_x_continuous(
    expand = c(0,0),
    limits = c(0,80),
    name = "age (years)") +
  coord_cartesian(clip = "off") +
  labs(
    caption = "Source: Encyclopedia Titanica",
    title = "Age distribution of passengers on the Titanic") +
  theme(
    panel.background = element_blank(),
    panel.grid.major.y = element_line(color = "#cbcbcb", size = 0.5),
    axis.ticks = element_blank(),
    axis.text = element_text(family = "Avenir Next", color = "#353d42", size = 10),
    axis.title = element_text(family = "Avenir Next", color = "#353d42", size = 11),
    plot.title = element_text(family = "Avenir Next", size = 13, color = "#353d42", face = "bold", margin = margin(t = 20)),
    plot.caption = element_text(family = "Avenir Next", size = 10, color = "#666666", hjust = 0)
  )
```

### Two densities

```{r warning = FALSE, message = FALSE, out.width='100%', fig.asp= 0.75, fig.align='center', echo = FALSE}
titanic2 <- titanic
titanic2$Sex <- factor(titanic2$Sex, levels = c("female", "male"))

ggplot(titanic2, aes(x = Age, y = ..count.., fill = Sex, color = Sex)) +
  geom_density_line(alpha = 0.6, bw = 2) +
  scale_y_continuous(
    expand = c(0,0),
    limits = c(0, 22),
    name = "scaled density") +
  scale_x_continuous(
    expand = c(0,0),
    limits = c(0,80),
    name = "age (years)") +
  scale_fill_manual(values = c("#0072B2", "#D55E00"), name = "gender") +
  scale_color_manual(values = c("darkblue","darkorange"), name = "gender")+
  guides(fill = guide_legend(override.aes = list(linetype = 0))) +
  coord_cartesian(clip = "off") +
  labs(
    caption = "Source: Encyclopedia Titanica",
    title = "Age distribution of passengers on the Titanic") +
  theme(
    panel.background = element_blank(),
    panel.grid.major.y = element_line(color = "#cbcbcb", size = 0.5),
    axis.ticks = element_blank(),
    axis.text = element_text(family = "Avenir Next", color = "#353d42", size = 11),
    axis.title = element_text(family = "Avenir Next", color = "#353d42", size = 11),
    plot.title = element_text(family = "Avenir Next", size = 13, color = "#353d42", face = "bold", margin = margin(b = 20)),
    plot.caption = element_text(family = "Avenir Next", size = 10, color = "#353d42", hjust = 0),
    legend.position = c(0.9, 0.8),
    legend.justification = c("right", "top"),
    legend.text = element_text(family = "Avenir Next", color = "#353d42", size = 11),
    legend.title = element_text(family = "Avenir Next", color = "#353d42", size = 11)
  )
```

```{r out.width='100%', fig.asp= 0.70, fig.align='center', echo = FALSE}

ggplot(titanic2, aes(x = Age, y = ..count..)) +
  geom_density_line(
    data = dplyr::select(titanic, -Sex),
    aes(fill = "all passengers"),
    color = "transparent") +
  geom_density_line(
    aes(fill = Sex), 
    bw = 2, 
    color = "transparent") +
  scale_y_continuous(
    expand = c(0,0),
    limits = c(0, 30),
    name = "scaled density") +
  scale_x_continuous(
    expand = c(0,0),
    limits = c(0,80),
    name = "age (years)") +
  scale_fill_manual(
    values = c("#b3b3b3a0", "#0072B2", "#D55E00"),
    breaks = c("all passengers", "male", "female"),
    labels = c("all passeengers", "males", "females"),
    name = NULL,
    guide = guide_legend(direction = "horizontal")
    ) +
  coord_cartesian(clip = "off") +
  facet_wrap(~Sex, labeller = labeller(Sex = function(Sex) paste(Sex, "passengers"))) +
  labs(
    caption = "Source: Encyclopedia Titanica",
    title = "Age distribution of passengers on the Titanic") +
  theme(
    panel.background = element_blank(),
    panel.grid.major.y = element_line(color = "#cbcbcb", size = 0.5),
    axis.ticks = element_blank(),
    axis.text = element_text(family = "Avenir Next", color = "#353d42", size = 10),
    axis.title = element_text(family = "Avenir Next", color = "#353d42", size = 11),
    strip.text = element_text(size = 11),
    strip.background = element_blank(),
    plot.title = element_text(family = "Avenir Next", size = 13, color = "#353d42", face = "bold", margin = margin(b = 10)),
    plot.caption = element_text(family = "Avenir Next", size = 11, color = "#666666", hjust = 0),
    legend.position = "bottom",
    legend.justification = "right",
    legend.text = element_text(family = "Avenir Next", color = "#353d42", size = 10),
    legend.spacing.x = grid::unit(4.5, "pt"),
    panel.spacing = grid::unit(1, "lines")
  )
```

### Multiple densities

```{r warning = FALSE, message = FALSE,  results = FALSE}
Cows %>% 
  mutate (breed = as.character(breed)) %>% 
  filter(breed != "Canadian") -> cows_filtered

# compute densities for sepal lengths
cows_dens <- group_by(cows_filtered, breed) %>% 
  do(ggplot2:::compute_density(.$butterfat, NULL)) %>% 
  rename(butterfat = x)

# get the maximum values
cows_max <-filter(cows_dens, density == max(density)) %>% 
  ungroup() %>% 
  mutate(
    hjust = c(0, 0, 0, 0),
    vjust = c(0, 0, 0, 0),
    nudge_x = c(-0.2, -0.2, 0.1, 0.23),
    nudge_y = c(0.03, 0.03, -0.2, -0.06))
```

```{r out.width='100%', fig.asp= 0.70, fig.align='center', echo = FALSE}

ggplot(cows_dens, aes(x = butterfat, y = density, color = breed, fill = breed)) + 
  geom_density_line(stat = "identity") +
  geom_text(
    data = cows_max,
    aes(
      label = breed, hjust = hjust, vjust = vjust,
      color = breed,
      x = butterfat + nudge_x, 
      y = density + nudge_y,
      family = "Avenir Next",
      size = 11
    ),
    inherit.aes = FALSE,
    size = 12/.pt
  ) +
  scale_color_manual(
    values = darken(c("#56B4E9", "#E69F00", "#D55E00", "#009E73"), 0.3),
    breaks = c("Ayrshire", "Guernsey", "Holstein-Friesian", "Jersey"),
    guide = "none"
  ) +
  scale_fill_manual(
    values = c("#56B4E950", "#E69F0050", "#D55E0050", "#009E7350"),
    breaks = c("Ayrshire", "Guernsey", "Holstein-Friesian", "Jersey"),
    guide = "none"
  ) +
  scale_x_continuous(
    expand = c(0, 0),
    labels = scales::percent_format(accuracy = 1, scale = 1),
    name = "butterfat contents"
  ) +
  scale_y_continuous(limits = c(0, 1.99), expand = c(0, 0)) +
  coord_cartesian(clip = "off") +
  labs(
    caption = "Source: Canadian Record of Performance for Purebred Dairy Cattle",
    title = "Butterfat percentage in the milk of four cattle breeds") +
  theme(
    panel.background = element_blank(),
    panel.grid.major.y = element_line(color = "#cbcbcb", size = 0.5),
    axis.ticks = element_blank(),
    axis.text = element_text(family = "Avenir Next", color = "#353d42", size = 11),
    axis.title = element_text(family = "Avenir Next", color = "#353d42", size = 11),
    plot.title = element_text(family = "Avenir Next", size = 13, color = "#353d42", face = "bold", margin = margin(b = 10)),
    plot.caption = element_text(family = "Avenir Next", size = 10, color = "#666666", hjust = 0))
```

# Visualize correlation
## Scatterplot
### Basic scatterplot
```{r out.width='100%', fig.asp= 0.8, fig.align='center', echo = FALSE}
data(BlueJays)
ggplot (BlueJays, aes(x = Mass, y = Head, fill = KnownSex)) + 
  geom_point(pch = 21, color = "white", size = 2.5) +
  scale_x_continuous(name = "body mass (g)") +
  scale_y_continuous(name = "head length (mm)") +
  scale_fill_manual(
    values = c(F = "#D55E00", M = "#0072B2"),
    breaks = c("F", "M"),
    labels = c("female birds", "male birds"),
    name = NULL,
    guide = guide_legend(
      direction = "horizontal",
      override.aes = list(size = 3))) +
  labs(
    caption = "Source: Keith Tarvin, Oberlin College",
    title = "Head length versus body mass for 123 blue jays") +
  theme(
    axis.ticks = element_blank(),
    axis.text = element_text(family = "Avenir Next", size = 11, color = "#353d42"),
    axis.title = element_text(family = "Avenir Next", size = 11, color = "#353d42"),
    panel.background = element_rect(fill = "white"),
    panel.grid.major = element_line(color = "#cbcbcb", size = 0.3),
    axis.title.x = element_text(margin = margin (t = 10)),
    plot.title = element_text(family = "Avenir Next", color = "#353d42", size = 15, face = "bold", margin = margin(b = 20)),
    plot.caption = element_text(family = "Avenir Next", color = "#666666", size = 9, hjust = 0, vjust = 2),
    legend.position = "top",
    legend.justification = "right",
    legend.box.spacing = unit(3.5, "pt"), #distance between legend and plot
    legend.text = element_text(family = "Avenir Next", color = "#353d42", vjust = 0.6),
    legend.spacing.x = unit(2, "pt"),
    legend.background = element_rect(fill = "white", color = "white"),
    legend.key.width = unit(10, "pt"),
    legend.key = element_blank())
```

### All-against-all scatterplot
```{r warning = FALSE, message = FALSE,  results = FALSE}
bj_matrix <- BlueJays %>% 
  dplyr::select(BirdID, KnownSex, Head, Mass, Skull) %>% 
  gather(var_x, val_x, Head:Skull) %>% 
  left_join(dplyr::select(BlueJays, BirdID, Head, Mass, Skull)) %>% 
  gather(var_y, val_y, Head:Skull)
```

```{r warning = FALSE, message = FALSE, out.width='100%', fig.asp= 1.1, fig.align='center', echo = FALSE }
labels <- c(
  Head = "head length (mm)",
  Mass = "body mass (g)", 
  Skull = "skull size (mm)"
)

ggplot (bj_matrix, aes(x = val_x, y = val_y, fill = KnownSex)) + 
  geom_point(pch = 21, color = "white", size = 2, stroke = 0.2) +
  scale_x_continuous(
    expand_scale(mult = 0.1),
    breaks = scales::pretty_breaks(4, min.n = 3)) +
  scale_y_continuous(
    expand_scale(mult = 0.1),
    breaks = scales::pretty_breaks(4, min.n = 3)) +
  scale_fill_manual(
    values = c(F = "#D55E00", M = "#0072B2"),
    breaks = c("F", "M"),
    labels = c("female birds", "male birds"),
    name = NULL,
    guide = guide_legend(
      direction = "horizontal",
      override.aes = list(size = 2.5))) +
  labs(
    caption = "Source: Keith Tarvin, Oberlin College",
    title = "All-against-all scattereplot for 123 blue jays",
    x = NULL,
    y = NULL) +
  facet_grid(
    var_y ~ var_x,
    scales = "free",
    switch = "both", 
    labeller = labeller(
      var_x = labels, 
      var_y = labels
    )
  ) +
  coord_cartesian(clip = "off") +
  panel_border(colour = "grey85", size = 0.4) +
  theme(
    axis.ticks = element_blank(),
    axis.text = element_text(family = "Avenir Next", size = 10, color = "#353d42"),
    axis.title = element_blank(),
    panel.background = element_rect(fill = "white"),
    panel.grid.major = element_line(color = "#cbcbcb", size = 0.5),
    plot.title = element_text(family = "Avenir Next", color = "#353d42", size = 14, face = "bold", margin = margin(b = 10)),
    plot.caption = element_text(family = "Avenir Next", color = "#666666", size = 9, hjust = 0, vjust = 2, margin = margin(t = 20)),
    legend.position = "top",
    legend.justification = "right",
    legend.box.spacing = grid::unit(1, "pt"), #distance between legend and plot
    legend.text = element_text(family = "Avenir Next", color = "#353d42", vjust = 0.6),
    legend.spacing.x = unit(2, "pt"),
    legend.background = element_rect(fill = "white", color = "white"),
    legend.key.width = unit(10, "pt"),
    legend.key = element_blank(),
    strip.background = element_blank(),
    strip.placement = "outside",
    strip.text.x = element_text(family = "Avenir Next", color = "#353d42", vjust = 1, margin = margin(0,0,0,0)),
    strip.text.y = element_text(family = "Avenir Next", color = "#353d42", vjust = 0, angle = -90, margin = margin(0,3.5,0,0))
    )

```

## Bubble chart

```{r warning = FALSE, message = FALSE, out.width='100%', fig.asp= 0.6, fig.align='center', echo = FALSE}
BlueJays$sex <- ifelse(BlueJays$KnownSex == "F", "female birds", "male birds")
BlueJays$sex <- factor(BlueJays$sex, levels = c("female birds", "male birds"))

ggplot (BlueJays, aes(x = Mass, y = Head, fill = KnownSex, size = Skull)) + 
  geom_point(pch = 21, color = "white") +
  facet_wrap(~sex, ncol = 2, scales = "fixed")+
  scale_x_continuous(name = "body mass (g)") +
  scale_y_continuous(name = "head length (mm)", breaks = c(52, 54, 56, 58, 60)) +
  scale_fill_manual(
    values = c(F = "#D55E00", M = "#0072B2"),
    breaks = c("F", "M"),
    labels = c("female birds", "male birds"),
    name = NULL,
    guide = "none") + 
  scale_radius(
    name = "skull size (mm)",
    range = c(2, 7),
    limits = c(28, 34),
    breaks = c(28, 30, 32, 34),
    labels = c("28   ", "30   ", "32   ", "34"),
    guide = guide_legend(
      direction = "horizontal",
      title.position = "top",
      title.hjust = 0.5,
      label.position = "right",
      override.aes = list(fill = "gray40"))) +
  labs(
    caption = "Source: Keith Tarvin, Oberlin College",
    title = "Bubble chart for 123 blue jays",
    x = NULL,
    y = NULL) +
  theme(
    panel.background = element_rect(fill = "white"),
    panel.grid.major = element_line(color = "#cbcbcb", size = 0.5),
    axis.ticks = element_blank(),
    axis.text = element_text(family = "Avenir Next", size = 10, color = "#353d42"),
    axis.title = element_text(family = "Avenir Next", size = 10, color = "#353d42"),
    axis.title.x = element_text(margin = margin (t = 10)),
    legend.position = c(1, 0),
    legend.justification = c(1,0),
    legend.spacing.x = unit(2, "pt"),
    legend.spacing.y = unit(2, "pt"),
    legend.text = element_text(family = "Avenir Next", color = "#353d42", size = 10, vjust = 0.6),
    legend.title = element_text(family = "Avenir Next", color = "#353d42", size = 10),
    legend.background = element_rect(fill = "white", color = "white"),
    legend.key.width = unit(5, "pt"),
    legend.key = element_blank(),
    strip.text = element_text(family = "Avenir Next", color = "#353d42", size = 10, margin = margin(2, 0, 2, 0)),
    strip.background  = element_rect(
      fill = "grey85", colour = "grey85",
      linetype = 1, size = 0.25),
    plot.title = element_text(family = "Avenir Next", color = "#353d42", size = 14, face = "bold", margin = margin(b = 10)),
    plot.caption = element_text(family = "Avenir Next", color = "#666666", size = 9, hjust = 0, vjust = 2, margin = margin(t = 10)))

```

## Correlogram
```{r warning = FALSE, message = FALSE,  results = FALSE}
data <- cor(mtcars[,1:7])
corre <- melt(data)
```

```{r out.width='100%', fig.asp= 0.7, fig.align='center', echo = FALSE}
ggplot(filter(corre, as.integer(Var1) < as.integer(Var2)), aes(Var1, Var2, fill = value)) + 
  geom_tile(color = "white", size = 1) + 
  scale_x_discrete(position = "top", name = NULL, expand = c(0, 0)) +
  scale_y_discrete(name = NULL, expand = c(0, 0)) +
  scale_fill_continuous_divergingx(
    palette = "Tropic", rev = FALSE,
    limits = c(-0.9, 0.91),
    breaks = c(-0.9, 0, 0.9),
    labels = c("–0.9", "0", "0.9"),
    name = "correlation",
    guide = guide_colorbar(
      direction = "horizontal",
      label.position = "bottom",
      title.position = "top",
      barwidth = grid::unit(140, "pt"),
      barheight = grid::unit(17.5, "pt"),
      ticks.linewidth = 1
    )
  ) +
  coord_fixed() +
  labs(
    caption = "Source: mtcars",
    title = "Correlation in automobile aspects") +
  theme(
    panel.background = element_blank(),
    axis.text = element_text(family = "Avenir Next", color = "#353D42", size = 11),
    axis.line = element_blank(),
    axis.ticks = element_blank(),
    axis.ticks.length = grid::unit(3, "pt"),
    legend.position = c(.97, .0),
    legend.justification = c(1, 0),
    legend.title.align = 0.5,
    legend.text = element_text(family = "Avenir Next", color = "#353D42", size = 11),
    legend.title = element_text(family = "Avenir Next", color = "#353D42", size = 11),
    plot.title = element_text(family = "Avenir Next", size = 13, color = "#353d42", face = "bold", margin = margin(b = 10)),
    plot.caption = element_text(family = "Avenir Next", size = 10, color = "#666666", hjust = 0)
  )
```

```{r out.width='100%', fig.asp= 0.7, fig.align='center', echo = FALSE}

ggplot(filter(corre, as.integer(Var1) < as.integer(Var2)), aes(Var1, Var2, fill = value, size = abs(value))) + 
  geom_point(shape = 21, stroke = 0) + 
  scale_x_discrete(position = "top", name = NULL, expand = c(0, 0.5)) +
  scale_y_discrete(name = NULL, expand = c(0, 0.5)) +
  scale_size_area(max_size = 19, limits = c(0, 0.91), guide = "none")+
  scale_fill_continuous_divergingx(
    palette = "PuOr", rev = FALSE,
    limits = c(-0.9, 0.91),
    breaks = c(-0.9, 0, 0.9),
    labels = c("–0.9", "0", "0.9"),
    name = "correlation",
    guide = guide_colorbar(
      direction = "horizontal",
      label.position = "bottom",
      title.position = "top",
      barwidth = grid::unit(140, "pt"),
      barheight = grid::unit(17.5, "pt"),
      ticks.linewidth = 1
    )
  ) +
  coord_fixed() +
  labs(
    caption = "Source: mtcars",
    title = "Correlation in automobile aspects",
    subtitle = "The size of circles denotes the magnitude of each correlation") +
  theme(
    panel.background = element_blank(),
    axis.text = element_text(family = "Avenir Next", color = "#353D42", size = 11),
    axis.line = element_blank(),
    axis.ticks = element_blank(),
    axis.ticks.length = grid::unit(3, "pt"),
    legend.position = c(.97, .0),
    legend.justification = c(1, 0),
    legend.title.align = 0.5,
    legend.text = element_text(family = "Avenir Next", color = "#353D42", size = 11),
    legend.title = element_text(family = "Avenir Next", color = "#353D42", size = 11),
    plot.title = element_text(family = "Avenir Next", size = 13, color = "#353d42", face = "bold"),
    plot.subtitle = element_text(family = "Avenir Next", size = 11, color = "#353D42", hjust = 0),
    plot.caption = element_text(family = "Avenir Next", size = 10, color = "#666666", hjust = 0)
  )
```

## Slopegraph
```{r warning = FALSE, message = FALSE,  results = FALSE}
co2_emissions <- read.csv("/Users/huvi/Downloads/Project_website/Dataset/annual-co2-emissions-per-country.csv", sep= ",")

emissions_data <- co2_emissions %>%
  dplyr::select(country = Entity, year = Year, co2 = Annual.CO2.emissions) %>% 
  mutate(co2 = co2/1000000, year = as.numeric(year)) %>% 
  filter(
    country %in% c("Trinidad and Tobago", "Qatar", "United Arab Emirates", "Oman", "Bahrain", "Singapore", "Netherlands", "Kazakhstan", "Equatorial Guinea", "Kuwait"),
    year %in% c(2000, 2005, 2010))

labels <- 
  tibble(filter(emissions_data, year == 2010))
labels$nudge = c(.1, .1, .1, .1, .1, -4, .1, 3, 4, .1)

```

```{r warning = FALSE, message = FALSE, out.width='100%', fig.asp= 0.7, fig.align='center', echo = FALSE}

ggplot(filter(emissions_data, year != 2005), aes(x = year, y = co2)) +
  geom_line(aes(group = country), color = "gray60") +
  geom_point(color = "white", size = 3) +
  geom_point(color = "#0072B2", size = 2) +
  geom_text(
    data = labels,
    aes(
      x = 2010 + 0.45, 
      y = co2 + nudge,
      label = country
    ),
    family = "Avenir Next",
    size = 10/.pt,
    hjust = 0
  ) +
  scale_x_continuous(
    limits = c(2000, 2020),
    breaks = c(2000, 2010),
    labels = c("2000", "2010"),
    expand = expand_scale(add = c(1, 0)),
    name = NULL,
    position = "top"
  ) +
  scale_y_continuous(
    limits = c(-2, 260),
    expand = c(0, 0), 
    name = parse(text = "`CO`[2]*` emissions (tons / person)`")
  ) +
  labs(
      caption = "Source: World Bank",
      title = "CO2 emissions per persion in 2000 and 2010 for 10 countries") +
  theme(
    panel.background = element_blank(),
    axis.line.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.text.x = element_text(family = "Avenir Next", size = 11, color = "#353d42"),
    axis.title.y = element_text(family = "Avenir Next", size = 11, color = "#353d42"),
    axis.text.y = element_text(family = "Avenir Next", size = 11, color = "#353d42"),
    axis.line.y.left = element_line(color = "#353d42"),
    plot.title = element_text(family = "Avenir Next", size = 14, color = "#353d42", face = "bold", margin = margin(b = 10), hjust = -0.1 ),
      plot.caption = element_text(family = "Avenir Next", size = 10, color = "#666666", margin = margin(t = 10), hjust = -0.06)
  )
```

```{r warning = FALSE, message = FALSE, out.width='100%', fig.asp= 0.7, fig.align='center', echo = FALSE}

ggplot(emissions_data, aes(x = year, y = co2)) +
  geom_line(aes(group = country), color = "gray60") +
  geom_point(color = "white", size = 3) +
  geom_point(color = "#0072B2", size = 2) +
  geom_text(
    data = labels,
    aes(
      x = 2010 + 0.45, 
      y = co2 + nudge,
      label = country
    ),
    family = "Avenir Next",
    size = 10/.pt,
    hjust = 0
  ) +
  scale_x_continuous(
    limits = c(2000, 2020),
    breaks = c(2000, 2005, 2010),
    labels = c("2000", "2005", "2010"),
    expand = expand_scale(add = c(1, 0)),
    name = NULL,
    position = "top"
  ) +
  scale_y_continuous(
    limits = c(-2, 260),
    expand = c(0, 0), 
    name = parse(text = "`CO`[2]*` emissions (tons / person)`")
  ) +
  labs(
      #caption = "Source: World Bank",
      #title = "CO2 emissions per persion from 2000 to 2010 for 10 countries"
    ) +
  theme(
    panel.background = element_blank(),
    axis.line.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.text.x = element_text(family = "Avenir Next", size = 11, color = "#353d42"),
    axis.title.y = element_text(family = "Avenir Next", size = 11, color = "#353d42"),
    axis.text.y = element_text(family = "Avenir Next", size = 11, color = "#353d42"),
    axis.line.y.left = element_line(color = "#353d42"),
    plot.title = element_text(family = "Avenir Next", size = 14, color = "#353d42", face = "bold", margin = margin(b = 10), hjust = -0.2),
      plot.caption = element_text(family = "Avenir Next", size = 10, color = "#666666", margin = margin(t = 10), hjust = -0.06)
  )
```

# Visualization proportion
## Waffle chart
```{r out.width='100%', fig.asp= 0.5, fig.align='center', echo = FALSE}
data <- data.frame(
  names = c("China", "India", "USA", "Other countries"),
  vals = c(14, 14, 3, 43)
)
waffle(
  data, 
  rows = 6,
  colors = c("#8dd3c7", "#80b1d3", "#fdb462", "#fb8072"),
  xlab = "1 square = 10 million people",
  title = "Word population in 2021")
```

## Pie chart
```{r warning = FALSE, message = FALSE,  results = FALSE}
# Create Data
data <- data.frame(
  group=LETTERS[1:5],
  value=c(13,7,9,21,2)
)
data <- data %>% 
  arrange(value) %>% 
  mutate(
    total = sum(value),
    end_angle = 2*pi*cumsum(value)/total, #ending angle for each pie slice
    start_angle = lag(end_angle, default = 0), #starting angle for each pie slice
    mid_angle = 0.5*(start_angle + end_angle), #middle angle for each pie slice, for text label
    hjust = ifelse(mid_angle > pi, 1, 0),
    vjust = ifelse(mid_angle<pi/2 | mid_angle>3*pi/2, 0, 1)
  ) 
rpie = 1
rlabel = 1.05*rpie
```

```{r out.width='70%', fig.asp= 1, fig.align='center', echo = FALSE}
ggplot(data) +
  geom_arc_bar(
    aes(
      x0 = 0, y0 = 0, r0 = 0, r = rpie,
      start = start_angle, end = end_angle, fill = group
    ),
    color = "white",
    size = 0.5
  ) +
  geom_text( #label each pie
    aes(
      x = rlabel*sin(mid_angle),
      y = rlabel*cos(mid_angle),
      label = group,
      hjust = hjust, vjust = vjust
    ),
    family = "Avenir Next", size = 7
  ) +
  geom_text(#label value of each pie
    aes(
      x = 0.6*sin(mid_angle),
      y = 0.6*cos(mid_angle),
      label = value
    ),
    family = "Avenir Next", size = 7
  ) +
  coord_fixed(clip = "off") +
  scale_x_continuous(
    expand = c(0, 0),
    name = "",
    breaks = NULL,
    labels = NULL
  ) +
  scale_y_continuous(
    expand = c(0, 0),
    name = "",
    breaks = NULL,
    labels = NULL
  ) +
  scale_fill_manual(
    values = c("#7fc97f", "#beaed4", "#fdc086", "#ffff99", "#386cb0")
  ) +
  theme(
    panel.background = element_blank(),
    legend.position = "none",
    plot.margin = margin(t = 20, r = 20)
  )

```

## Donut chart
```{r out.width='70%', fig.asp= 1, fig.align='center', echo = FALSE}
data <- data.frame(
  category=c("A", "B", "C"),
  count=c(10, 60, 30)
)
# Compute percentages
data$fraction = data$count / sum(data$count)
# Compute the cumulative percentages (top of each rectangle)
data$ymax = cumsum(data$fraction)
# Compute the bottom of each rectangle
data$ymin = c(0, head(data$ymax, n=-1))
# Compute label position
data$labelPosition <- (data$ymax + data$ymin) / 2
# Compute a good label
data$label <- paste0(data$category, "\n value: ", data$count)

# Make the plot
ggplot(data, aes(ymax=ymax, ymin=ymin, xmax=4, xmin=3, fill=category)) +
  geom_rect() +
  coord_polar(theta="y") + # Try to remove that to understand how the chart is built 
  xlim(c(2, 4)) + # Try to remove that to see how to make a pie chart
  geom_label(x = 3.5, aes(y = labelPosition, label = label), size = 6, family = "Avenir Next") +
  scale_fill_manual(
    values = c("#66c2a5", "#fc8d62", "#8da0cb")
  ) +
  theme(
    panel.background = element_blank(),
    axis.ticks = element_blank(),
    axis.text = element_blank(),
    legend.position = "none",
    plot.margin = margin(t = 20, r = 20)
  )
```

## Mosaic plot

```{r warning = FALSE, message = FALSE,  results = FALSE}
colNames<-c("IDENTIF","RIVER","LOCATION","ERECTED","PURPOSE","LENGTH","LANES","CLEAR-G","T-OR-D","MATERIAL","SPAN","REL-L","TYPE")

bridge <- read.csv("/Users/huvi/Downloads/Project_website/Dataset/bridges.data.version.csv", col.names = colNames, na.strings=c("NA", "-", "?"," "))%>%
  dplyr::select(ERECTED, material = MATERIAL) %>%
  mutate(
    erected = case_when(
      ERECTED < 1870 ~ "craft",
      ERECTED >= 1870 & ERECTED <= 1889 ~ "emerging",
      ERECTED >= 1890 & ERECTED <= 1939 ~ "mature",
      ERECTED > 1940 ~ "modern"))

bridge %>%
  dplyr::select(material, erected) %>% 
  table() %>% 
  reshape2::melt() %>% 
  dplyr::select(material, erected, count = value) %>%
  mutate(
    material = case_when(
      material == "IRON" ~ "iron",
      material == "STEEL" ~ "steel",
      material == "WOOD" ~ "wood")
    ) %>%
  group_by(erected)%>%
  mutate(group_count = sum(count))-> bridges_tidy
  
labels_df <- group_by(bridges_tidy, erected) %>%
  filter(count != 0) %>%
  arrange(desc(material)) %>%
  mutate(
    y = (cumsum(count) - 0.5*count)/group_count,
    y = ifelse(
      erected == "mature" & material == "wood", NA, y
    )
  )
```

```{r warning = FALSE, message = FALSE, out.width='100%', fig.asp= 1, fig.align='center', echo = FALSE}

ggplot(bridges_tidy) +
  aes(x = erected, y = count, width = group_count, fill = material) +
  geom_bar(stat = "identity", position = "fill", colour = "white", size = 1) +
  facet_grid(~erected, scales = "free_x", space = "free_x") +
  geom_text(
    data = labels_df,
    aes(y = y, label = count, color = material),
    na.rm = TRUE,
    size = 12/.pt,
    family = "Avenir Next"
  ) +
  scale_y_continuous(
    name = NULL) +
  scale_x_discrete(
    name = NULL) +
  scale_fill_manual(
    values = c("#D55E00D0", "#0072B2D0", "#009E73D0")
  ) +
  scale_color_manual(
    values = c(iron = "white", wood = "white", steel = "white")
  ) +
  labs( title = "Mosaic plot for 106 bridges in Pittsburgh",
        subtitle = "The widths of each rectangle are proportional to the number of bridges built in that area
The heights of each rectangle are proportional to the number of bridges built from that material
Numbers show the counts of bridges in each category",
        caption = "Source: Yoram Reich and Steven J. Fenves") +
  coord_cartesian(clip = "off") +
  theme(
    panel.background = element_blank(),
    line = element_blank(),
    strip.text = element_blank(),
    axis.ticks.length = unit(0, "pt"),
    axis.text.y = element_blank(),
    axis.text.x = element_text(family = "Avenir Next", size = 11, color = "#353d42"),
    panel.spacing.x = unit(0, "pt"),
    plot.title = element_text(family = "Avenir Next", size = 15, color = "#353d42", face = "bold"),
    plot.subtitle = element_text(family = "Avenir Next", size = 11, color = "#353d42"),
    plot.caption = element_text(family = "Avenir Next", size = 10, color = "#666666", hjust = 0, margin = margin(t = 10)),
    legend.position = "top",
    legend.justification = c(0, 0),
    legend.title = element_blank(),
    legend.background = element_blank(),
    legend.key = element_blank(),
    legend.text = element_text(family = "Avenir Next", size = 11, color = "#353d42"),
    legend.margin = margin(b = -20, l = -2)
    )
```

## Treemap

```{r warning = FALSE, message = FALSE,  results = FALSE}
library(tidyverse)
library(tinter)
library(treemapify)

colNames<-c("IDENTIF","RIVER","LOCATION","ERECTED","PURPOSE","LENGTH","LANES","CLEAR-G","T-OR-D","MATERIAL","SPAN","REL-L","TYPE")

bridge <- read.csv("/Users/huvi/Downloads/Project_website/Dataset/bridges.data.version.csv", col.names = colNames, na.strings=c("NA", "-", "?"," "))%>%
  dplyr::select(ERECTED, material = MATERIAL) %>%
  mutate(
    erected = case_when(
      ERECTED < 1870 ~ "craft",
      ERECTED >= 1870 & ERECTED <= 1889 ~ "emerging",
      ERECTED >= 1890 & ERECTED <= 1939 ~ "mature",
      ERECTED > 1940 ~ "modern"))

bridge %>%
  dplyr::select(material, erected) %>% 
  table() %>% 
  reshape2::melt() %>% 
  dplyr::select(material, erected, count = value) %>%
  mutate(
    material = case_when(
      material == "IRON" ~ "iron",
      material == "STEEL" ~ "steel",
      material == "WOOD" ~ "wood")
    ) %>%
  group_by(erected)%>%
  mutate(group_count = sum(count))-> bridges_tidy
  
labels_df <- group_by(bridges_tidy, erected) %>%
  filter(count != 0) %>%
  arrange(desc(material)) %>%
  mutate(
    y = (cumsum(count) - 0.5*count)/group_count,
    y = ifelse(
      erected == "mature" & material == "wood", NA, y
    )
  )
```

```{r warning = FALSE, message = FALSE, out.width='100%', fig.asp= 1, fig.align='center', echo = FALSE}
filcols <- c("#D55E00D0", "#0072B2D0", "#009E73D0")
filcols <- c(vapply(filcols, function(x) c(lighten(x, .9), lighten(x, .6), lighten(x, .3), x), character(4)))

ggplot(bridges_tidy, aes(area = count, subgroup = material, fill = interaction(erected, material))) +
  geom_treemap(color = "white", size = 0.5*.pt, alpha = NA) + 
  geom_treemap_subgroup_text(
    family = "Avenir Next",
    colour = "grey50",
    place = "centre", alpha = 0.7,
    grow = TRUE
  ) +
  geom_treemap_subgroup_border(color = "white") +
  geom_treemap_text(
    aes(label = erected, color = interaction(erected, material)),
    family = "Avenir Next",
    place = "centre",
    grow = FALSE
  ) +
  scale_fill_manual(values = filcols) +
  scale_color_manual(values = c(
    crafts.iron = "black", crafts.steel = "black", crafts.wood = "black",
    emerging.iron = "black", emerging.steel = "black", emerging.wood = "black",
    mature.iron = "black", mature.steel = "black", mature.wood = "black",
    modern.iron = "white", modern.steel = "white", modern.wood = "white")
  ) +
  coord_cartesian(clip = "off") +
  guides(colour = "none", fill = "none")+
  labs( title = "Treemap plot for 106 bridges in Pittsburgh",
        subtitle = "The area of each rectangle is proportional to the number of bridges of that type",
        caption = "Source: Yoram Reich and Steven J. Fenves") +
  theme(
    plot.title = element_text(family = "Avenir Next", size = 15, color = "#353d42", face = "bold"),
    plot.subtitle = element_text(family = "Avenir Next", size = 11, color = "#353d42"),
    plot.caption = element_text(family = "Avenir Next", size = 10, color = "#666666", hjust = 0, margin = margin(t = 10))
    )
```

## Parallel sets plot
```{r warning = FALSE, message = FALSE,  results = FALSE}

colNames<-c("IDENTIF","RIVER","LOCATION","ERECTED","PURPOSE","LENGTH","LANES","CLEAR-G","T-OR-D","MATERIAL","SPAN","REL-L","TYPE")

bridge <- read.csv("/Users/huvi/Downloads/Project_website/Dataset/bridges.data.version.csv", col.names = colNames, na.strings=c("NA", "-", "?"," "))%>%
  mutate(
    erected = case_when(
      ERECTED < 1870 ~ "CRAFTS",
      ERECTED >= 1870 & ERECTED <= 1889 ~ "EMERGING",
      ERECTED >= 1890 & ERECTED <= 1939 ~ "MATURE",
      ERECTED > 1940 ~ "MODERN"),
    length = case_when(
      LENGTH < 1500 ~ "SHORT",
      ERECTED >= 1500 & ERECTED <= 3000 ~ "MEDIUM",
      ERECTED > 3000 ~ "LONG"))

bridge %>% 
  dplyr::select(material = MATERIAL, erected, river = RIVER, length) %>% 
  filter(river != "Y") %>% 
  table() %>% 
  reshape2::melt() %>% 
  rename(count = value) %>%
  mutate(
    material = factor(
      case_when(
        material == "IRON" ~ "iron",
        material == "STEEL" ~ "steel",
        material == "WOOD" ~ "wood"
      ),
      levels = c("wood", "steel", "iron")
    ),
    erected = factor(
      case_when(
        erected == "CRAFTS" ~ "crafts",
        erected == "EMERGING" ~ "emerging",
        erected == "MATURE" ~ "mature",
        erected == "MODERN" ~ "modern"
      ),
      levels = c("modern", "mature", "emerging", "crafts")
    ),
    length = factor(
      case_when(
        length == "LONG" ~ "long",
        length == "MEDIUM" ~ "medium",
        length == "SHORT" ~ "short"
      ),
      levels = c("short", "medium", "long")
    ),
    river = factor(
      case_when(
        river == "A" ~ "Allegheny",
        river == "M" ~ "Monongahela",
        river == "O" ~ "Ohio"
      ),
      levels = c("Ohio", "Monongahela", "Allegheny")
    )
  ) -> data

data <- gather_set_data(data, 1:4)
data$x <- factor(data$x, levels = c("material", "length", "erected", "river"))
```

```{r out.width='100%', fig.asp= 0.8, fig.align='center', echo = FALSE}
ggplot(data, aes(x, id = id, split = y, value = count)) +
  geom_parallel_sets(aes(fill = material), alpha = 0.5, axis.width = 0.13) +
  geom_parallel_sets_axes(axis.width = 0.1, fill = "grey80", color = "grey80") +
  geom_parallel_sets_labels(
    color = 'black',
    family = "Avenir Next",
    size = 10/.pt,
    angle = 90
  ) +
  scale_x_discrete(
    name = NULL,
    expand = c(0, 0.2)
  ) +
  scale_y_continuous(breaks = NULL, expand = c(0, 0))+
  scale_fill_manual(
    values = c(iron = "#D55E00D0", wood = "#009E73D0", steel = "#0072B2D0"),
    guide = "none"
  ) +
  labs( title = "Parallel sets plot for 106 bridges in Pittsburgh",
        subtitle = "The coloring of the bands highlights the construction material of different bridges",
        caption = "Source: Yoram Reich and Steven J. Fenves") +
  theme(
    panel.background = element_blank(),
    axis.line = element_blank(),
    axis.ticks = element_blank(),
    axis.text = element_text(family = "Avenir Next", size = 11, color = "#353d42"),
    strip.text = element_text(family = "Avenir Next", size = 11, color = "#353d42"),
    plot.margin = margin(14, 1.5, 2, 1.5),
    plot.title = element_text(family = "Avenir Next", size = 15, color = "#353d42", face = "bold", hjust = 0.13),
    plot.subtitle = element_text(family = "Avenir Next", size = 11, color = "#353d42", hjust = 0.22),
    plot.caption = element_text(family = "Avenir Next", size = 10, color = "#666666", margin = margin(t = 10, b = 5), hjust = 0.05)
  )
```

# Visualize revolution
## Line chart
### Single variable
```{r warning = FALSE, message = FALSE, out.width='100%', fig.asp= 0.8, fig.align='center', echo = FALSE}
covid <- read.csv("/Users/huvi/Downloads/Project_website/Dataset/owid-covid-data.csv") 

ger <- covid %>% filter(iso_code == "DEU") %>% 
  dplyr::select(date, new_cases_per_million) %>% 
  mutate(
    date = as.Date(date),
    MA7 = zoo::rollmean(new_cases_per_million, k = 7, fill = NA))

ggplot(ger, aes(x = date, y = MA7))+ 
  geom_line(color = "#0072B2", size = 0.6)+
  scale_y_continuous(
    limit = c(0, 350),
    expand = c(0, 0),
    breaks = c(0, 50, 100, 150, 200, 250, 300),
    name = "cases/million people"
  ) +
  scale_x_date(
    name = "date"
  ) +
  labs(
    caption = "Source: ourworldindata.com",
    title = "7-day moving average of new COVID cases in Germany") +
  theme(
    axis.text = element_text(family = "Avenir Next", size = 11, color = "#353d42"),
    axis.title = element_text(family = "Avenir Next", size = 11, color = "#353d42"),
    axis.line = element_line(color = "#353d42", size = 0.4),
    panel.background = element_blank(),
    panel.grid.major = element_blank(),
    axis.title.x = element_text(margin = margin (t = 10)),
    plot.title = element_text(family = "Avenir Next", color = "#353d42", size = 15, face = "bold", margin = margin(b = 20)),
    plot.caption = element_text(family = "Avenir Next", color = "#666666", size = 9, hjust = 0, vjust = 2),
  )

```

```{r warning = FALSE, message = FALSE, out.width='100%', fig.asp= 0.8, fig.align='center', echo = FALSE}

vaccine <- covid %>% filter(location %in% c("United Kingdom", "Germany", "Israel", "United States", "United Arab Emirates", "Canada" )) %>% 
  dplyr::select(date, location, total_vaccinations_per_hundred) %>% 
  mutate(
    date = as.Date(date)
    ) %>% 
  drop_na()

```

### Multiple variables
```{r warning = FALSE, message = FALSE, out.width='100%', fig.asp= 0.8, fig.align='center', echo = FALSE}

ggplot(vaccine, aes(x = date, y = total_vaccinations_per_hundred, color = location))+ 
  geom_line(size = 0.6)+
  scale_y_continuous(
    limit = c(0, 125),
    expand = c(0, 0),
    breaks = c(0, 25, 50, 75, 100, 125),
    name = "vaccinations/hundred people"
  ) +
  scale_x_date(
    name = "date"
  ) +
  scale_color_manual(
    values = c("#e41a1c", "#377eb8", "#4daf4a", "#984ea3", "#ff7f00", "#6a3d9a")
  ) +
  labs(
    caption = "Source: ourworldindata.com",
    title = "Number of vaccinations per hundred people in 6 countries") +
  theme(
    axis.text = element_text(family = "Avenir Next", size = 11, color = "#353d42"),
    axis.title = element_text(family = "Avenir Next", size = 11, color = "#353d42"),
    axis.line = element_line(color = "#353d42", size = 0.4),
    panel.background = element_blank(),
    panel.grid.major = element_blank(),
    axis.title.x = element_text(margin = margin (t = 10)),
    plot.title = element_text(family = "Avenir Next", color = "#353d42", size = 15, face = "bold", margin = margin(b = 20)),
    plot.caption = element_text(family = "Avenir Next", color = "#666666", size = 9, hjust = 0, vjust = 2),
    legend.position = "top",
    legend.box.spacing = unit(3.5, "pt"), #distance between legend and plot
    legend.text = element_text(family = "Avenir Next", color = "#353d42", vjust = 0.6),
    legend.spacing.x = unit(2, "pt"),
    legend.background = element_rect(fill = "white", color = "white"),
    legend.key.width = unit(10, "pt"),
    legend.key = element_blank(),
    legend.title = element_blank()
    )
```

## Area chart
```{r warning = FALSE, message = FALSE, out.width='100%', fig.asp= 0.8, fig.align='center', echo = FALSE}

ggplot(ger, aes(x = date, height = MA7, y = 0))+ 
  geom_ridgeline(color = "#0072B2", fill = "#0072B240", size = 0.75)+
  scale_y_continuous(
    limit = c(0, 350),
    expand = c(0, 0),
    breaks = c(0, 50, 100, 150, 200, 250, 300),
    name = "cases/million people"
  ) +
  scale_x_date(
    name = "date"
  ) +
  labs(
    caption = "Source: ourworldindata.com",
    title = "7-day moving average of new COVID cases in Germany") +
  theme(
    axis.text = element_text(family = "Avenir Next", size = 11, color = "#353d42"),
    axis.title = element_text(family = "Avenir Next", size = 11, color = "#353d42"),
    axis.line = element_line(color = "#353d42", size = 0.4),
    panel.background = element_blank(),
    panel.grid.major = element_blank(),
    axis.title.x = element_text(margin = margin (t = 10)),
    plot.title = element_text(family = "Avenir Next", color = "#353d42", size = 15, face = "bold", margin = margin(b = 20)),
    plot.caption = element_text(family = "Avenir Next", color = "#666666", size = 9, hjust = 0, vjust = 2),
  )

```