

# Clean the data
```{python}
import pandas as pd

aapl = (pd.read_csv("/Users/huvi611/Downloads/HistoricalData_1671538564695.csv")
          .assign(Date = lambda x: pd.to_datetime(x.Date),
                  Open = lambda x: x.Open.str.replace('$','', regex = True).astype('float'),
                  High = lambda x: x.High.str.replace('$','', regex = True).astype('float'),
                  Low = lambda x: x.Low.str.replace('$','', regex = True).astype('float'),
                  Close = lambda x: x["Close/Last"].str.replace('$','', regex = True).astype('float'),
          )
        )  

pd.read_csv("/Users/huvi611/Downloads/HistoricalData_1671538564695.csv")
```

# Visualize the data

The visualization shows changes in the price and volume of Apple stock between 2005 and 2021. The top line chart represents the price, while the bottom column chart represents the volume. I also included important milestones and historical events that affected the stock.

The dataset comes from [The Nasdaq Stock Market](https://www.nasdaq.com/market-activity/stocks/aapl/historical).


```{r message = FALSE, warning = FALSE, out.width = '100%', fig.align='center', fig.asp = 0.5, eval = FALSE, results = FALSE}
library(tidyverse)
library(lubridate)
library(cowplot)

text_color = "#353d42"
caption_color = "#666666"
font = "Avenir Next"

aapl <- read.csv("/Users/huvi611/Downloads/HistoricalData_1671538564695.csv")%>% 
  select(date = Date, price = Open, volume = Volume) %>%
  mutate(
    date = ymd(date)
  )



price <- ggplot(aapl, aes(x = date, y = price)) +
  #global financial crisis  
  annotate("rect", xmin = as.Date("2007-12-01"), xmax = as.Date("2009-06-01"), ymin=0, ymax=Inf, alpha=0.2, fill="#cbcbcb") +
  annotate(geom="text", x = as.Date("2008-09-01"), y = 90, label = "Global\nfinancial crisis", family = font, color = text_color, fontface = "bold")+
  
  #Covid  
  annotate("rect", xmin = as.Date("2019-12-01"), xmax = as.Date("2021-05-01"), ymin=0, ymax=Inf, alpha=0.2, fill="#cbcbcb") +
  annotate(geom="text", x = as.Date("2020-08-01"), y = 30, label = "COVID-19\npandemic", family = font, color = text_color, fontface = "bold")+  

  #10/01/2006: first macbookpro
  annotate(geom="text", x = as.Date("2006-01-10"), y = 12, label = "First\nMacBook Pro", family = font, color = text_color, fontface = "bold")+
  geom_segment(x = as.Date("2006-01-10"), xend = as.Date("2006-01-10"), y = 4, yend = 7, color = text_color) +
    
  #09/01/2007: first AppleTV
  annotate(geom="text", x = as.Date("2007-01-09"), y = 29, label = "First\nApple TV", family = font, color = text_color, fontface = "bold")+
  geom_segment(x = as.Date("2007-01-09"), xend = as.Date("2007-01-09"), y = 4, yend = 25, color = text_color) +
    
  #29/06/2007: first iphone
  annotate(geom="text", x = as.Date("2007-06-29"), y = 15, label = "First\niPhone", family = font, color = text_color, fontface = "bold")+
  geom_segment(x = as.Date("2007-06-29"), xend = as.Date("2007-06-29"), y = 6, yend = 10, color = text_color) +
    
  #27/01/2010: first ipad
  annotate(geom="text", x = as.Date("2010-01-27"), y = 17, label = "First iPad", family = font, color = text_color, fontface = "bold")+
  geom_segment(x = as.Date("2010-01-27"), xend = as.Date("2010-01-27"), y = 8, yend = 15, color = text_color) +
    
  #24/08/2011: Job resigns
  annotate(geom="text", x = as.Date("2011-08-24"), y = 26, label = "Steve Jobs\nresigned", family = font, color = text_color, fontface = "bold")+
  geom_segment(x = as.Date("2011-08-24"), xend = as.Date("2011-08-24"), y = 14, yend = 21, color = text_color) +
    
  #12/12/2012: iphone 5
  annotate(geom="text", x = as.Date("2012-09-17"), y = 37, label = "iPhone5", family = font, color = text_color, fontface = "bold")+
  annotate(geom="point", x=as.Date("2012-09-17"), y = 25.5, size = 3, shape=21, fill="transparent", col = text_color) +
  geom_segment(x = as.Date("2012-09-17"), xend = as.Date("2012-09-17"), y = 27, yend = 35, color = text_color) +
  #12/12/2012: iphone XS
  annotate(geom="text", x = as.Date("2018-09-21"), y = 70, label = "iPhone XS", family = font, color = text_color, fontface = "bold")+
  annotate(geom="point", x=as.Date("2018-09-21"), y = 58, size = 3, shape=21, fill="transparent", col = text_color) +
  geom_segment(x = as.Date("2018-09-21"), xend = as.Date("2018-09-21"), y = 59, yend = 68, color = text_color) + 
    
  #24/04/2015: first apple watch
  annotate(geom="text", x = as.Date("2015-04-24"), y = 45, label = "First \nApple watch", family = font, color = text_color, fontface = "bold")+
  geom_segment(x = as.Date("2015-04-24"), xend = as.Date("2015-04-24"), y = 34, yend = 41, color = text_color) +
    
  #03/11/2017: iphoneX
  annotate(geom="text", x = as.Date("2017-11-03"), y = 95, label = "iPhone X", family = font, color = text_color, fontface = "bold")+
  geom_segment(x = as.Date("2017-11-03"), xend = as.Date("2017-11-03"), y = 44, yend = 90, color = text_color) +
  
  #25/03/2019: Apple TV+
  annotate(geom="text", x = as.Date("2019-03-25"), y = 104, label = "Apple TV+\nApple News+\nApple Arcade\nApple Card", family = font, color = text_color, fontface = "bold")+
  geom_segment(x = as.Date("2019-03-25"), xend = as.Date("2019-03-25"), y = 50, yend = 95, color = text_color) +  
  
  #20/09/2019: iphone11
  annotate(geom="text", x = as.Date("2019-09-20"), y = 85, label = "iPhone 11", family = font, color = text_color, fontface = "bold")+
  geom_segment(x = as.Date("2019-09-20"), xend = as.Date("2019-09-20"), y = 58, yend = 83, color = text_color) + 
    
  #24/04/2020: iphonese
  annotate(geom="text", x = as.Date("2020-02-20"), y = 126, label = "iPhone SE", family = font, color = text_color, fontface = "bold")+
  geom_segment(x = as.Date("2020-04-24"), xend = as.Date("2020-04-24"), y = 75, yend = 124, color = text_color) + 
    
  #26/01/2021: all-time high
  annotate(geom="text", x = as.Date("2020-7-01"), y = 152.16, label = "All-time high \n$143.16", family = font, color = text_color, fontface = "bold") +
  annotate(geom="point", x=as.Date("2021-01-26"), y=143.5, size = 3, shape=21, fill="transparent", col = text_color) +
  geom_segment(x = as.Date("2021-01-26"), xend = as.Date("2021-01-26"), y = 145, yend = 155, color = text_color) 


,
    panel.grid.major = element_line(color = "grey90", size = 0.3, linetype = "dashed"),
    axis.ticks.y = element_blank(),
    axis.ticks.x = element_line(color = "#grey90", size = 0.5),
    axis.text = element_text(family = "Open Sans", size = 11, color = "grey50", face = "bold")
```

```{r}
# price

ggplot(py$aapl, aes(x = Date)) +
  geom_line(aes(y = Open), size = 0.5, color = "#69b3a2")+ 
  scale_y_continuous(expand = c(0,0),
                     name = NULL,
                     limits = c(0, 350),
                     breaks = seq(0,160,20),
                     labels = paste("$",seq(0,160,20))) +                     
  scale_x_date(
    expand = c(0,0),
    breaks = as.Date(c("2006-01-01", "2007-01-01", "2008-01-01", "2009-01-01","2010-01-01","2011-01-01", "2012-01-01","2013-01-01", "2014-01-01","2015-01-01", "2016-01-01", "2017-01-01","2018-01-01", "2019-01-01","2020-01-01","2021-01-01", "2021-01-01")),
    labels = seq(2006,2022, 1),
    name = NULL) +
  theme(
    panel.background = element_rect(color = "white", fill = "white"),
    panel.grid.major = element_line(color = "grey90", size = 0.3, linetype = "dashed"), 
    axis.ticks.x = element_line(color = "grey90", size = 0.5))
```


```{r}
# volume

b = ggplot(py$aapl, aes(x = Date, y = Volume)) +
  geom_col(color="#69b3a2")+
  scale_y_reverse(
    expand = c(0,0),
    name = NULL) +
  scale_x_date(
    expand = c(0,0),
    breaks = as.Date(c("2006-01-01", "2007-01-01", "2008-01-01", "2009-01-01","2010-01-01","2011-01-01", "2012-01-01","2013-01-01", "2014-01-01","2015-01-01", "2016-01-01", "2017-01-01","2018-01-01", "2019-01-01","2020-01-01","2021-01-01", "2021-01-01")),
    labels = seq(2006,2022, 1),
    name = NULL) +
  theme(
    panel.background = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks = element_blank(),
    axis.text = element_blank()
  )
```

```{r}
a + b + plot_layout(ncol = 1, heights = c(3, 1))
ggsave("apple.png", width = 10, height = 10)
```



- Add apple logo
- Compare with other big tech companies
- Build a shiny app



