---
title: "Data_viz_1_astronaut"
output: html_document
date: "2022-11-26"
---


<font size ="5"><center> **2020 Daily Temperatures in Kiel, Germany**</center></font>

----

The two heatmaps show the daily maximum and minimum temperatures observed in Holtenau station in Kiel, Germany. All observations were record by [Deutscher Wetterdienst](www.dwd.de) in the period between 01.01.2020 and 31.01.2020

**How to read the heatmap** 

- The x-axis is a discrete scale that represents days of the month
- The y-axis is a discrete scale that represents months of the year 
- The color represents daily temperatures. I use the Oranges and Blues sequential color scale from [ColorBrewer](https://colorbrewer2.org/)) for maximum and minimum temperatures, respectively. 

```{r message = FALSE, warning = FALSE, out.width = '100%', fig.align='center', fig.asp=0.6}
library(dplyr)
library(ggplot2)
library(lubridate)
text_color = "#353d42"
caption_color = "#666666"
font = "Avenir Next"
kiel_tem <- read.csv("/Users/huvi/Downloads/kiel_temperature.csv") %>% 
  mutate(
    tem_max = round((TMAX - 32)*5/9,1),
    tem_min = round((TMIN - 32)*5/9,1) 
  ) %>% 
  mutate(
    date = ymd(DATE),
    year = year(date),
    month = as.factor(month(date, label = TRUE)),
    day = as.factor(day(date))
  )%>% 
  select(month, day, tem_max, tem_min)
ggplot(kiel_tem, aes(x = day, y = month, fill = tem_max)) +
  geom_tile(color = "#f0f0f0", size = 0.5) + #Color size of border between tiles
  coord_equal() + # draw square instead of rectangular tiles
  scale_x_discrete(name = NULL, position = "top") +
  scale_y_discrete(expand = c(0, 0), name = NULL) +
  scale_fill_distiller(
    type = "seq",
    palette = "Oranges",
    direction = 1,
    limits = c(1, 32),
    name = "max temperature (°C)",
    guide = guide_colorbar(
      barwidth = grid::unit(2, "in"),
      barheight = grid::unit(0.2, "in"),
      title.position = "top",
      title.hjust = 0.5
    )
  ) +
  theme(
    panel.background = element_blank(),
    axis.ticks = element_blank(),
    axis.text = element_text(family = font, color = text_color, size = 7),
    legend.title = element_text(family = font, color = text_color, size = 7),
    legend.text = element_text(family = font, color = text_color, size = 7),
    legend.position = "top",
    legend.justification = "right",
    plot.margin = margin(t = 0, b = 0, r = 0, l = 0)
  )
```

```{r message = FALSE, warning = FALSE, out.width = '100%', fig.align='center', fig.asp=0.6}
ggplot(kiel_tem, aes(x = day, y = month, fill = tem_min)) +
  geom_tile(color = "#f0f0f0", size = 0.5) + #Color size of border between tiles
  coord_equal() + # draw square instead of rectangular tiles
  scale_x_discrete(name = NULL, position = "top") +
  scale_y_discrete(expand = c(0, 0), name = NULL) +
  scale_fill_distiller(
    type = "div",
    palette = "Blues",
    direction = 1,
    limits = c(-3, 19),
    name = "min temperature (°C)",
    guide = guide_colorbar(
      barwidth = grid::unit(2, "in"),
      barheight = grid::unit(0.2, "in"),
      title.position = "top",
      title.hjust = 0.5
    )
  ) +
  theme(
    panel.background = element_blank(),
    axis.ticks = element_blank(),
    axis.text = element_text(family = font, color = text_color, size = 7),
    legend.title = element_text(family = font, color = text_color, size = 7),
    legend.text = element_text(family = font, color = text_color, size = 7),
    legend.position = "top",
    legend.justification = "right",
    plot.margin = margin(t = 0, b = 0, r = 0, l = 0)
  )
```


**Some highlights**

- August is the hottest month of the year. Temperatures are high between June and September
- Low-temperatures months include December, January, February, and March
- Maximum temperatures range from 1°C to 31°C
- Minimum temperatures range from -3°C to 19°C

Personal opinion: Compared to other parts of Germany, the temperature range in Kiel is quite temperature. The weather is quite nice, but sometimes it can also be too windy (for a bike rider like me :D) 










